---
title: "All data clustering"
output: html_notebook
---

```{r Setup}
library(tidyverse)
library(ggplot2)
library(caret)
library(factoextra)
library(corrr)
library(ggcorrplot)
library(FactoMineR)
library(doParallel)
```

# use more cpu cores
```{r}
registerDoParallel(cores = detectCores() - 1)
```

# read data
```{r}
data <- read_rds("../data/data_aggregated.rds")
head(data)
data2 <-  data %>%
  select(-partition_date, -id, -vehicleid,-timestamp, -ignitionState, -event_description, -terminal_id, -recieved_ts, -gps_fix_type,   -gps_accuracy, -SP_CODE,-SP_NAME, -MP_NAME, -MN_NAME, -DC_NAME, -PR_NAME, -created_ts, -country_name)
head(data2)
```

# normalizing
```{r}
data_norm <- scale(data2)
```

# principle component analysis
## inspecting pairwise combinations
```{r}
pairs(data_norm)
```

## inspecting correlation
```{r}
data_norm_corr_matrix <- cor(data_norm)
ggcorrplot(data_norm_corr_matrix)
```

## inspecting covariance
```{r}
cov(data_norm)
```

## importance of components
```{r}
data_norm_pca <- princomp(data_norm_corr_matrix)
summary(data_norm_pca)
```

## relation of components to columns
```{r}
data_norm_pca$loadings[, 1:2]
```

# visualizations
## importance of principle components
```{r}
fviz_eig(data_norm_pca, addlabels = TRUE)
```

## principle components similarities and dissimilarities, attribute impact on principle components
```{r}
fviz_pca_var(data_norm_pca, col.var = "black")
```

## contribution of attributes to principle components
```{r}
fviz_cos2(data_norm_pca, choice = "var", axes = 1:2)
```

### optimum clusters
```{r}
factoextra::fviz_nbclust(x=data,
                         kmeans,
                         method = c("wss"))

factoextra::fviz_nbclust(x=data_norm,
                         kmeans,
                         method = c("silhouette"))
```


```{r}
cluster_model <- kmeans(data_norm,
                        centers = 3,
                        nstart = 10,
                        iter.max = 25)
```

