---
title: "Untitled"
output: html_document
date: "2023-08-17"
---

```{r}
library(tidyverse)
library(ggplot2)
library(caret)
library(factoextra)
```


#Reading in both the claims data and the taxi data
```{r}
claims <- read.csv("../data/claims_data.csv")
taxis <- read_rds("../data/all_data.rds")
claims
```
#filtering the claims data to only the taxis that have submitted claims
```{r}
claimed <- claims %>% 
  filter(number.of.claims > 0) %>% 
  filter(vehicleid != 155517428)
claimed
```
#removing rows where the gps is not accurate, joining the two tables
```{r}
claimed2 <- claimed %>% 
  inner_join(taxis, claimed2, by = "vehicleid") %>% 
  filter(gps_fix_type == 3)
claimed2
```
#removing the taxis that have not laid a claim, and filtering down the data to the useful columns
```{r}
oneclaim <- claimed2 %>% 
  filter(number.of.claims > 0) %>% 
  select(vehicleid, timestamp, speed, linear_g, lateral_g)
oneclaim
```
#getting the greaters loinear acceleration in a negetive direction for each of the taxis
```{r}
minLinear <- oneclaim %>% 
  group_by(vehicleid) %>% 
  summarize(linear_g = min(linear_g)) %>% 
  left_join(oneclaim, minLinear, by = "linear_g") %>% 
  arrange(linear_g) %>% 
  filter(vehicleid.x == vehicleid.y) %>% 
  rename(vehicleid = "vehicleid.x") %>% 
  select(!"vehicleid.y")
minLinear
```
#getting the maximum lateral g foirce for each of the taxis
```{r}
maxLat <- oneclaim %>% 
  group_by(vehicleid) %>% 
  summarize(lateral_g = max(lateral_g))%>% 
  left_join(oneclaim, maxLat, by = "lateral_g") %>% 
  arrange(desc(lateral_g)) %>% 
  filter(vehicleid.x == vehicleid.y) %>% 
  rename(vehicleid = "vehicleid.x") %>% 
  select(!"vehicleid.y")
maxLat
```
#getting the greatesty linear acceleration in the positive direction for each taxi
```{r}
maxLinear <- oneclaim %>% 
  group_by(vehicleid) %>% 
  summarize(linear_g = max(linear_g)) %>% 
  left_join(oneclaim, minLinear, by = "linear_g") %>% 
  arrange(desc(linear_g)) %>% 
  filter(vehicleid.x == vehicleid.y) %>% 
  rename(vehicleid = "vehicleid.x") %>% 
  select(!"vehicleid.y")
maxLinear
```
#looking at the specific timestamps for each event for each taxi
```{r}
joinTable <- minLinear %>% 
  left_join(maxLat, minLinear, by = "vehicleid") %>% 
  rename(Minlinear_g = "linear_g.x", MinLinSpeed = "speed.x", MinLinLateral_g = "lateral_g.x", MinLinTimestamp = "timestamp.x") %>% 
  rename(MaxLateral_g = "lateral_g.y", MaxLatTimestamp = "timestamp.y", maxLatSpeed = "speed.y", maxLatLinear_g = "linear_g.y") %>% 
  left_join(maxLinear, joinTable, by = "vehicleid") %>% 
  rename(maxLinear_g = "linear_g", maxlinTimestamp = "timestamp", maxlinspeed = "speed", maxLinLateral_g = "lateral_g") %>% 
  distinct(vehicleid, MinLinTimestamp, MaxLatTimestamp,  maxlinTimestamp)
joinTable
```
#creating a list of times where the max and mins are closly togeter indication an accident on the road
```{r}
joinTableTimes <- joinTable %>% 
  select(vehicleid, MinLinTimestamp, maxlinTimestamp,  MaxLatTimestamp) %>% 
  mutate(linearDifference = MinLinTimestamp - maxlinTimestamp) %>% 
  mutate(MinLinearLateralDifference = MinLinTimestamp - MaxLatTimestamp) %>% 
  mutate(MaxLinearLateralDifference = maxlinTimestamp - MaxLatTimestamp) %>% 
  filter((linearDifference < 300 & linearDifference > -300) | (MinLinearLateralDifference < 5 & MinLinearLateralDifference >-5) | (MaxLinearLateralDifference < 0.833 & MaxLinearLateralDifference >-0.833))
joinTableTimes
```
#Combining the above table with the claimed table to see the number of claims for each taxi in the list
```{r}
joinTablesClaims <- joinTableTimes %>% 
  left_join(claimed, joinTableTimes, by = "vehicleid") %>% 
  mutate(accident = TRUE)
joinTablesClaims
```
#combineing table and classifying risk based on accidents
```{r}
accidentRisk <- claims %>% 
  select(vehicleid) %>% 
  left_join(joinTablesClaims, claims, by = "vehicleid") %>% 
  select(vehicleid, accident) %>% 
  mutate(accident = ifelse(is.na(accident),
                           FALSE,
                           TRUE)) %>% 
  mutate(risk = ifelse(accident == TRUE, 
         10, 
         1)) %>% 
  select(!accident)
  length(unique(accidentRisk$`vehicleid`))

accidentRisk
```
```{r}
costRisk <- claims %>% 
  #replacing NULL with 0 
  mutate(total.claims.cost = ifelse(total.claims.cost == "NULL", 0, total.claims.cost)) %>%
  #converting total.claims.cost from char to double
  mutate(total.claims.cost = as.double(total.claims.cost)) %>% 
  #classifying cost risk
  mutate(costRisk = case_when(total.claims.cost > 250000 ~ 10,
                               total.claims.cost > 125000 ~ 7,
                               total.claims.cost > 100000 ~ 6,
                               total.claims.cost > 85000 ~ 5,
                               total.claims.cost > 60000 ~ 4,
                               total.claims.cost > 40000 ~ 3,
                               total.claims.cost < 40000 & total.claims.cost > 0  ~ 2,
                              total.claims.cost == 0 ~ 1)) %>%
  #classifyling claim risk
  mutate(claimRisk = case_when(number.of.claims > 4 ~ 10,
                               number.of.claims > 2 ~ 5,
                               number.of.claims > 0 ~ 3,
                               number.of.claims == 0 ~ 1)) %>% 
  #calculating cost per claim risk
  mutate(costPerClaimRisk = case_when((total.claims.cost/number.of.claims) > 150000 ~ 10,
                                      (total.claims.cost/number.of.claims) > 120000 ~ 7,
                                      (total.claims.cost/number.of.claims) > 80000 ~ 5,
                                      (total.claims.cost/number.of.claims) > 60000 ~ 4,
                                      (total.claims.cost/number.of.claims) > 30000 ~ 3,
                                      (total.claims.cost/number.of.claims) > 0 ~ 2,
                                      (total.claims.cost+number.of.claims) == 0 ~ 1,)) %>% 
  group_by(vehicleid) %>% 
  mutate(totalClaimsRisk = (costRisk * 0.50 + claimRisk * 0.30 + costPerClaimRisk * 0.20)) %>% 
  ungroup(vehicleid) %>% 
  arrange(desc(totalClaimsRisk))
  
costRisk
```
```{r}
claimsRiskFinal <- costRisk %>% 
  select(vehicleid, totalClaimsRisk)
claimsRiskFinal

saveRDS(claimsRiskFinal, file = "../data/claimsRiskFinal.rds")
```


