---
title: "Untitled"
output: html_document
date: "2023-08-17"
---

```{r}
library(tidyverse)
library(ggplot2)
library(caret)
library(factoextra)
```


#Reading in both the claims data and the taxi data
```{r}
claims <- read.csv("claims_data (1).csv")
taxis <- read_rds("all_data.rds")
claims
```
#filtering the claims data to only the taxis that have submitted claims
```{r}
claimed <- claims %>% 
  filter(number.of.claims > 0) %>% 
  filter(vehicleid != 155517428)
claimed
```
#removing rows where the gps is not accurate, joining the two tables
```{r}
claimed2 <- claimed %>% 
  inner_join(taxis, claimed2, by = "vehicleid") %>% 
  filter(gps_fix_type == 3)
claimed2
```
#removing the taxis that have not laid a claim, and filtering down the data to the useful columns
```{r}
oneclaim <- claimed2 %>% 
  filter(number.of.claims > 0) %>% 
  select(vehicleid, timestamp, speed, linear_g, lateral_g)
oneclaim
```
#getting the greaters loinear acceleration in a negetive direction for each of the taxis
```{r}
minLinear <- oneclaim %>% 
  group_by(vehicleid) %>% 
  summarize(linear_g = min(linear_g)) %>% 
  left_join(oneclaim, minLinear, by = "linear_g") %>% 
  arrange(linear_g) %>% 
  filter(vehicleid.x == vehicleid.y) %>% 
  rename(vehicleid = "vehicleid.x") %>% 
  select(!"vehicleid.y")
minLinear
```
#getting the maximum lateral g foirce for each of the taxis
```{r}
maxLat <- oneclaim %>% 
  group_by(vehicleid) %>% 
  summarize(lateral_g = max(lateral_g))%>% 
  left_join(oneclaim, maxLat, by = "lateral_g") %>% 
  arrange(desc(lateral_g)) %>% 
  filter(vehicleid.x == vehicleid.y) %>% 
  rename(vehicleid = "vehicleid.x") %>% 
  select(!"vehicleid.y")
maxLat
```
#getting the greatesty linear acceleration in the positive direction for each taxi
```{r}
maxLinear <- oneclaim %>% 
  group_by(vehicleid) %>% 
  summarize(linear_g = max(linear_g)) %>% 
  left_join(oneclaim, minLinear, by = "linear_g") %>% 
  arrange(desc(linear_g)) %>% 
  filter(vehicleid.x == vehicleid.y) %>% 
  rename(vehicleid = "vehicleid.x") %>% 
  select(!"vehicleid.y")
maxLinear
```
#looking at the specific timestamps for each event for each taxi
```{r}
joinTable <- minLinear %>% 
  left_join(maxLat, minLinear, by = "vehicleid") %>% 
  rename(Minlinear_g = "linear_g.x", MinLinSpeed = "speed.x", MinLinLateral_g = "lateral_g.x", MinLinTimestamp = "timestamp.x") %>% 
  rename(MaxLateral_g = "lateral_g.y", MaxLatTimestamp = "timestamp.y", maxLatSpeed = "speed.y", maxLatLinear_g = "linear_g.y") %>% 
  left_join(maxLinear, joinTable, by = "vehicleid") %>% 
  rename(maxLinear_g = "linear_g", maxlinTimestamp = "timestamp", maxlinspeed = "speed", maxLinLateral_g = "lateral_g") %>% 
  distinct(vehicleid, MinLinTimestamp, MaxLatTimestamp,  maxlinTimestamp)
joinTable
```
#creating a list of times where the max and mins are closly togeter indication an accident on the road
```{r}
joinTableTimes <- joinTable %>% 
  select(vehicleid, MinLinTimestamp, maxlinTimestamp,  MaxLatTimestamp) %>% 
  mutate(linearDifference = MinLinTimestamp - maxlinTimestamp) %>% 
  mutate(MinLinearLateralDifference = MinLinTimestamp - MaxLatTimestamp) %>% 
  mutate(MaxLinearLateralDifference = maxlinTimestamp - MaxLatTimestamp) %>% 
  filter((linearDifference < 300 & linearDifference > -300) | (MinLinearLateralDifference < 5 & MinLinearLateralDifference >-5) | (MaxLinearLateralDifference < 0.833 & MaxLinearLateralDifference >-0.833))
joinTableTimes
```
#Combining the above table with the claimed table to see the number of claims for each taxi in the list
```{r}
joinTablesClaims <- joinTableTimes %>% 
  left_join(claimed, joinTableTimes, by = "vehicleidA")
joinTablesClaims
```

