---
title: "TaxiCluster"
output: html_document
date: "2023-08-15"
---
# Clustering 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(caret)
library(factoextra)
library(corrr)
library(ggcorrplot)
library(FactoMineR)
library(ksharp)
library(ppclust)
library(fclust)
library(cluster)

```

## reading in data

```{r}
averageTimeOn <- read_rds("../data/average_speed_intervals.rds")
taxiAnalysis <- read_rds("../data/TaxiRiskAnalysis.rds")
data_shift <- read_rds("../data/shift_risk.rds") %>% 
  drop_na()
averageTimeOn
taxiAnalysis
```

## preprocessing
### data selection
getting the time spent on from the table by removing the time spent off from the dataset, also selecting the relevant columns
```{r}
averageTimeOnTotal <- averageTimeOn %>% 
  dplyr::select(vehicleid, time_ON_Sec, distance_KM, average_speed_KMpH,event_description) %>% 
  filter(event_description == "Ign OFF")
  

averageTimeOnTotal
```

### taxi averages
averaging the data out to be used for clustering analysis

```{r}
averageTimeOnTotalAvg <- averageTimeOnTotal %>% 
  group_by(vehicleid) %>% 
  summarise(time_ON_SecAvg = mean(as.numeric(time_ON_Sec)), 
            distance_KMAvg = mean(distance_KM), 
            average_speed_KMpHAvg = mean(average_speed_KMpH)) 

averageTimeOnTotalAvg
```

### join operation
joining the two tables to allow for clustering 
- removed vehicle id's
```{r}
taxiAnalysis <- right_join(data_shift,taxiAnalysis, by = "vehicleid" )

clusterTable <- taxiAnalysis %>% 
  left_join(averageTimeOnTotalAvg, taxiAnalysis, by = "vehicleid") 
clusterTable

#averaging N/As
clusterTable <- clusterTable %>% 
  mutate_if(is.numeric, ~replace_na(.,mean(., na.rm = TRUE)))

```

### normalizing
```{r}
clusterTable_normalized <- clusterTable
```

## principle component analysis
### inspecting pairwise combinations
```{r}
pairs(clusterTable_normalized)
```

### inspecting correlation
```{r}
clusterTable_corr_matrix <- cor(clusterTable_normalized)
ggcorrplot(clusterTable_corr_matrix)

```

### inspecting covariance
```{r}

cov(clusterTable_normalized)
```

### importance of components
```{r}
clusterTable_pca <- princomp(clusterTable_corr_matrix)
summary(clusterTable_pca)
```

### relation of components to columns
```{r}
clusterTable_pca$loadings[, 1:2]
```

## visualizations
### importance of principle components
```{r}
fviz_eig(clusterTable_pca, addlabels = TRUE)
```

### principle components similarities and dissimilarities, attribute impact on principle components
```{r}
fviz_pca_var(clusterTable_pca, col.var = "black")
```
variables grouped tightly together are positively related, opposite variables are negatively related, and the distance from origin is how well represented the variable is.

### contribution of attributes to principle components
```{r}
fviz_cos2(clusterTable_pca, choice = "var", axes = 1:2)
```

### optimum clusters
```{r}
factoextra::fviz_nbclust(x=clusterTable_normalized,
                         kmeans,
                         method = c("wss"))

factoextra::fviz_nbclust(x=clusterTable_normalized,
                         kmeans,
                         method = c("silhouette"))
```

## modeling optimum clusters
```{r}

cluster_model <- kmeans(clusterTable_normalized,
                        centers = 3,
                        nstart = 5,
                        iter.max = 10)
cluster_model
```

## appending data with relevant clusters
```{r}
taxiRiskClustered <- cbind(clusterTable_normalized,
                           cluster = cluster_model$cluster)
head(taxiRiskClustered)
```

# visualizing clusters
```{r}
fviz_cluster(cluster_model, clusterTable_normalized, 
             ellipse.type = "convex",
             geom=c("point"), 
             palette = "jco", 
             ggtheme = theme_classic(),
             repel = TRUE) 

```
# overlapping clusters - more than one driver

```{r}
res.fcm <- fcm(clusterTable, centers=4, nstart=5)
output_df <- as.data.frame(res.fcm$u)[1:104,]
vehicleid <- clusterTable$vehicleid
membership_df <- as.data.frame(res.fcm$u)
combined_df <- cbind(vehicleid, membership_df, output_df)
summary(combined_df)
res.fcm$v

```

```{r}

newData <- combined_df[,1:4]
options(scipen = 999)

overlapping_taxi <- newData %>% 
  mutate(HasOverlap = case_when(`Cluster 1` >= 0.2 & `Cluster 2` >= 0.2 ~ "HasOverlap",
                                `Cluster 1` >= 0.2 & `Cluster 3` >= 0.2 ~ "HasOverlap",
                                `Cluster 2` >= 0.2 & `Cluster 3` >= 0.2 ~ "HasOverlap",
                                TRUE ~ "No_Overlap"))
overlapping_taxi

taxisWithOverlap <- overlapping_taxi %>% 
  filter(HasOverlap == "HasOverlap")

clusterTable_overlap <- clusterTable %>% 
  filter(vehicleid %in% c(111082600,115994042,119059287,122129229,122196196))

write.csv(clusterTable_overlap,file = "../data/multiple_drivers_risk_info.csv")
saveRDS(clusterTable_overlap, file = "../data/multiple_drivers_risk_info.rds")

```

