---
title: "taxihack"
output: html_document
date: "2023-08-11"
---


```{r}
library(tidyverse)
library(ggplot2)
library(caret)
library(factoextra)
```
#reading in the claims data and the data for the 1st taxi in the list

```{r}
#firstTaxi <- read.csv("104804470.csv")#When running with all the data comment out this line
#claimsData <- read.csv("claims_data (1).csv")
```
#line to be used when running with all data
#name is due to how the origional code was written and im too lazy to fix it 
```{r}
firstTaxi <- read_rds("all_data.rds")
```
#starting to look at the speed of the taxi, as well as the locations where the taxi speeds
#could use this data as part of the cluster analysis or to compare for profiles portentually, have not done this for the rest but is easy to do - thoughts?
```{r}
firstTaxiSpeed <- firstTaxi %>% 
  select("partition_date", "vehicleid", "timestamp", "speed", "road_speed", "SP_NAME") %>% 
  mutate(road_speed = as.integer(road_speed)) %>% 
  mutate(timestamp = as.character(timestamp)) %>% 
  mutate(vehicleid = as.integer(vehicleid)) %>% 
  mutate(speed = as.integer(speed)) %>% 
  mutate(partition_date = as.character(partition_date))
firstTaxiSpeed
```
#Looking at the percentage of time the taxi is speeding in each of the areas - again this has not been done for the other factors but is doable
```{r}
firstTaxiSpeedingPercentage <- firstTaxiSpeed %>% 
  drop_na(road_speed) %>% 
  group_by(SP_NAME, vehicleid) %>% 
  count(speed > road_speed) %>% 
  pivot_wider(names_from = "speed > road_speed", values_from = n) %>% 
  rename(Slower = "FALSE") %>% 
  rename(Faster = "TRUE") %>% 
  mutate(Slower = replace_na(Slower, 0)) %>% 
  mutate(Faster = replace_na(Faster, 0)) %>%
  mutate(total = Slower + Faster) %>% 
  mutate(speedingPercentage = Faster/total * 100)
  
firstTaxiSpeedingPercentage
```
#Looking at the times when the taxi is at a finable speed (10km or more over the speed limit), some of these are gps failures though - reccomendation to "board" is that this is taken with a pinch of salt

#Example looking at the source data, the event at '2022-01-02 19:28:38' with vehicleid = 104804470 is a result of the gps picking the taxi up on the wrong road.
```{r}
firstTaxiSpeeding10More <- firstTaxiSpeed %>% 
  drop_na(road_speed) %>% 
  group_by(SP_NAME, vehicleid) %>% 
  filter((speed - road_speed) >= 10) %>% 
  mutate(difference = speed - road_speed) %>% 
  arrange(desc(difference)) 
firstTaxiSpeeding10More
```
#Looking at which area the taxi is speeding on average in, it is possible to ensure that the taxis that speed on average avoid these areas
```{r}
firstTaxiSpeedAvg <- firstTaxiSpeed %>% 
  select(vehicleid, SP_NAME, speed, road_speed) %>% 
  group_by(SP_NAME, vehicleid) %>% 
  summarise(mean(road_speed), mean(speed)) %>% 
  rename(avgspeed = "mean(speed)") %>% 
  rename(avglimit = "mean(road_speed)") %>% 
  mutate(Speeding = case_when(avgspeed > avglimit ~ "TRUE",
                              avgspeed <= avglimit ~ "FALSE"))
firstTaxiSpeedAvg
```
#a condensed list of areas where specific taxis speed on average
```{r}
firstTaxiSpeedAvgnot <- firstTaxiSpeedAvg %>% 
  filter(Speeding == TRUE)
firstTaxiSpeedAvgnot
```
#shortening the dataset for ease of use for risk analysis
```{r}
firstTaxiSpeedAvg1 <- firstTaxiSpeedAvg %>% 
  mutate(avgDifference = avglimit - avgspeed) %>% 
  select(SP_NAME, avgDifference)
firstTaxiSpeedAvg1
```
summary of the table above
```{r}
summary(firstTaxiSpeedAvg1)
```
#Assigning a risk value to each average speed difference of each taxi in each area

#Taxi A in area X has a speed risk rating of Q
#10 is worst, 10 is best
```{r}
firstTaxiSpeedRisk <- firstTaxiSpeedAvg %>% 
  mutate(avgDifference = avglimit - avgspeed) %>% 
  mutate(speedRiskLevel = case_when(
                               avgDifference < summary(firstTaxiSpeedAvg1$`avgDifference`)[1] + ((summary(firstTaxiSpeedAvg1$`avgDifference`)[6] - summary(firstTaxiSpeedAvg1$`avgDifference`)[1])*0.1) & avgDifference > 0 ~ "10",
                               avgDifference < summary(firstTaxiSpeedAvg1$`avgDifference`)[1] + ((summary(firstTaxiSpeedAvg1$`avgDifference`)[6] - summary(firstTaxiSpeedAvg1$`avgDifference`)[1])*0.2) & avgDifference > 0 ~ "9",
                               avgDifference < summary(firstTaxiSpeedAvg1$`avgDifference`)[1] + ((summary(firstTaxiSpeedAvg1$`avgDifference`)[6] - summary(firstTaxiSpeedAvg1$`avgDifference`)[1])*0.3) & avgDifference > 0 ~ "8",
                               avgDifference < summary(firstTaxiSpeedAvg1$`avgDifference`)[1] + ((summary(firstTaxiSpeedAvg1$`avgDifference`)[6] - summary(firstTaxiSpeedAvg1$`avgDifference`)[1])*0.4) & avgDifference > 0 ~ "7",
                               avgDifference < summary(firstTaxiSpeedAvg1$`avgDifference`)[1] + ((summary(firstTaxiSpeedAvg1$`avgDifference`)[6] - summary(firstTaxiSpeedAvg1$`avgDifference`)[1])*0.5) & avgDifference > 0 ~ "6",
                               avgDifference < summary(firstTaxiSpeedAvg1$`avgDifference`)[1] + ((summary(firstTaxiSpeedAvg1$`avgDifference`)[6] - summary(firstTaxiSpeedAvg1$`avgDifference`)[1])*0.6) & avgDifference > 0 ~ "5",
                               avgDifference < summary(firstTaxiSpeedAvg1$`avgDifference`)[1] + ((summary(firstTaxiSpeedAvg1$`avgDifference`)[6] - summary(firstTaxiSpeedAvg1$`avgDifference`)[1])*0.7) & avgDifference > 0 ~ "4",
                               avgDifference < summary(firstTaxiSpeedAvg1$`avgDifference`)[1] + ((summary(firstTaxiSpeedAvg1$`avgDifference`)[6] - summary(firstTaxiSpeedAvg1$`avgDifference`)[1])*0.8) & avgDifference > 0 ~ "3",
                               avgDifference < summary(firstTaxiSpeedAvg1$`avgDifference`)[1] + ((summary(firstTaxiSpeedAvg1$`avgDifference`)[6] - summary(firstTaxiSpeedAvg1$`avgDifference`)[1])*0.9) & avgDifference > 0 ~ "2",
                               avgDifference < summary(firstTaxiSpeedAvg1$`avgDifference`)[1] + ((summary(firstTaxiSpeedAvg1$`avgDifference`)[6] - summary(firstTaxiSpeedAvg1$`avgDifference`)[1])*1.0) & avgDifference > 0 ~ "1",
                               avgDifference <= 0 ~ "10")) %>% 
  arrange(as.numeric(speedRiskLevel))
firstTaxiSpeedRisk
```
#Getting the average speed risk rateing for each driver
```{r}
firstTaxiSpeedRiskAvg <- firstTaxiSpeedRisk %>% 
  group_by(vehicleid) %>% 
  drop_na(speedRiskLevel) %>% 
  summarise(avgSpeedRisk = mean(as.numeric(speedRiskLevel)))
firstTaxiSpeedRiskAvg
```
#Starting to look at the cornering of each taxi
```{r}
firstTaxiCorner <- firstTaxi%>% 
  select("partition_date", "vehicleid", "timestamp","lateral_g", "SP_NAME") %>% 
  filter(lateral_g > 0.1 & lateral_g < 0.42) 
firstTaxiCorner
```
```{r}
summary(firstTaxiCorner)
```

#Jumping straight to assigning a risk value to each cornering of each taxi
 
#Cornering at more than 0.42 gs is considered dangerous as most cars will start to slide and even flip, thus a rating of 10 is given to any lateral g greater than 0.42
```{r}
firstTaxiCornerRisk <- firstTaxiCorner %>% 
  mutate(cornerRisk = case_when(
                   lateral_g < summary(firstTaxiCorner$`lateral_g`)[1] + ((summary(firstTaxiCorner$`lateral_g`)[6] - summary(firstTaxiCorner$`lateral_g`)[1])*0.1)~ "1",
                   lateral_g < summary(firstTaxiCorner$`lateral_g`)[1] + ((summary(firstTaxiCorner$`lateral_g`)[6] - summary(firstTaxiCorner$`lateral_g`)[1])*0.2)~ "2",
                   lateral_g < summary(firstTaxiCorner$`lateral_g`)[1] + ((summary(firstTaxiCorner$`lateral_g`)[6] - summary(firstTaxiCorner$`lateral_g`)[1])*0.3)~ "3",
                   lateral_g < summary(firstTaxiCorner$`lateral_g`)[1] + ((summary(firstTaxiCorner$`lateral_g`)[6] - summary(firstTaxiCorner$`lateral_g`)[1])*0.4)~ "4",
                   lateral_g < summary(firstTaxiCorner$`lateral_g`)[1] + ((summary(firstTaxiCorner$`lateral_g`)[6] - summary(firstTaxiCorner$`lateral_g`)[1])*0.5)~ "5",
                   lateral_g < summary(firstTaxiCorner$`lateral_g`)[1] + ((summary(firstTaxiCorner$`lateral_g`)[6] - summary(firstTaxiCorner$`lateral_g`)[1])*0.6)~ "6",
                   lateral_g < summary(firstTaxiCorner$`lateral_g`)[1] + ((summary(firstTaxiCorner$`lateral_g`)[6] - summary(firstTaxiCorner$`lateral_g`)[1])*0.7)~ "7",
                   lateral_g < summary(firstTaxiCorner$`lateral_g`)[1] + ((summary(firstTaxiCorner$`lateral_g`)[6] - summary(firstTaxiCorner$`lateral_g`)[1])*0.8)~ "8",
                   lateral_g < summary(firstTaxiCorner$`lateral_g`)[1] + ((summary(firstTaxiCorner$`lateral_g`)[6] - summary(firstTaxiCorner$`lateral_g`)[1])*0.9)~ "9",
                   lateral_g < summary(firstTaxiCorner$`lateral_g`)[1] + ((summary(firstTaxiCorner$`lateral_g`)[6] - summary(firstTaxiCorner$`lateral_g`)[1])*01.0)~ "10")) %>% 
  arrange(desc(as.numeric(cornerRisk)))
firstTaxiCornerRisk
```
#Getting the average lateral g force for each taxi
```{r}
firstTaxiCornerAvg <- firstTaxiCornerRisk %>% 
  group_by(vehicleid) %>% 
  drop_na(cornerRisk) %>% 
  summarise(avgLateral_g = mean(lateral_g)) 
firstTaxiCornerAvg

```
#getting the cornering risk average for each taxi
```{r}
firstTaxiCornerRiskAvg <- firstTaxiCornerRisk %>% 
  group_by(vehicleid) %>% 
  drop_na(cornerRisk) %>% 
  summarise(avgconerRisk = mean(as.numeric(cornerRisk)))
firstTaxiCornerRiskAvg
```
#combining the above two tables
```{r}
firstTaxiCornerCombine <- firstTaxiCornerAvg %>% 
  left_join(firstTaxiCornerRiskAvg, firstTaxiCornerAvg, by = c("vehicleid"))
firstTaxiCornerCombine
```
#starting to look at the each braking event of each taxi
```{r}
firstTaxiBraking <- firstTaxi %>% 
  select(partition_date, vehicleid, linear_g, timestamp) %>% 
  filter(linear_g < -0.2 & linear_g > -0.47) %>% 
  arrange(desc(linear_g))
firstTaxiBraking
```
#getting the summary of the braking 
```{r}
summary(firstTaxiBraking)
```
#assigning the a risk rating to each braking event for each taxi

#Breaking harder than -0.5 gs is considered dangerous and thus scores lower than -0.5 are given a 10 rating for risk
```{r}
firstTaxiBrakingRisk <- firstTaxiBraking %>% 
  mutate(brakingRisk = case_when(linear_g < summary(firstTaxiBraking$`linear_g`)[1] + ((summary(firstTaxiBraking$`linear_g`)[6] - summary(firstTaxiBraking$`linear_g`)[1])*0.1)~ "10",
                   linear_g < summary(firstTaxiBraking$`linear_g`)[1] + ((summary(firstTaxiBraking$`linear_g`)[6] - summary(firstTaxiBraking$`linear_g`)[1])*0.2)~ "9",
                   linear_g < summary(firstTaxiBraking$`linear_g`)[1] + ((summary(firstTaxiBraking$`linear_g`)[6] - summary(firstTaxiBraking$`linear_g`)[1])*0.3)~ "8",
                   linear_g < summary(firstTaxiBraking$`linear_g`)[1] + ((summary(firstTaxiBraking$`linear_g`)[6] - summary(firstTaxiBraking$`linear_g`)[1])*0.4)~ "7",
                   linear_g < summary(firstTaxiBraking$`linear_g`)[1] + ((summary(firstTaxiBraking$`linear_g`)[6] - summary(firstTaxiBraking$`linear_g`)[1])*0.5)~ "6",
                   linear_g < summary(firstTaxiBraking$`linear_g`)[1] + ((summary(firstTaxiBraking$`linear_g`)[6] - summary(firstTaxiBraking$`linear_g`)[1])*0.6)~ "5",
                   linear_g < summary(firstTaxiBraking$`linear_g`)[1] + ((summary(firstTaxiBraking$`linear_g`)[6] - summary(firstTaxiBraking$`linear_g`)[1])*0.7)~ "4",
                   linear_g < summary(firstTaxiBraking$`linear_g`)[1] + ((summary(firstTaxiBraking$`linear_g`)[6] - summary(firstTaxiBraking$`linear_g`)[1])*0.8)~ "3",
                   linear_g < summary(firstTaxiBraking$`linear_g`)[1] + ((summary(firstTaxiBraking$`linear_g`)[6] - summary(firstTaxiBraking$`linear_g`)[1])*0.9)~ "2",
                   linear_g < summary(firstTaxiBraking$`linear_g`)[1] + ((summary(firstTaxiBraking$`linear_g`)[6] - summary(firstTaxiBraking$`linear_g`)[1])*01.0)~ "1")) %>% 
  arrange(desc(as.numeric(brakingRisk)))
firstTaxiBrakingRisk
```
#getting the average risk rating for breaking for each taxi
```{r}
firstTaxiBrakingRiskAvg <- firstTaxiBrakingRisk %>% 
  group_by(vehicleid) %>% 
  drop_na(brakingRisk) %>% 
  summarise(avgBrakingRisk = mean(as.numeric(brakingRisk)))
firstTaxiBrakingRiskAvg
```
#startingthe analysis of acceleration
```{r}
firstTaxiAcc <- firstTaxi %>% 
  select(partition_date, vehicleid, linear_g, timestamp) %>% 
  filter(linear_g > 0.2 & linear_g < 0.47) %>% 
  arrange(desc(linear_g))
firstTaxiAcc
```
#summary of the acceleration table
```{r}
summary(firstTaxiAcc)
```
#assigning a risk rating to the acceleration of the taxis 

#accelerating harder than 0.4 gs is bad for the vehicle, can cause discomfort for riders and could be portentually dangerous and thus all scores greater 0.4 are given a 10 rating
```{r}
firstTaxiAccRisk <- firstTaxiAcc %>% 
  mutate(accelerationRisk = case_when(linear_g < summary(firstTaxiAcc$`linear_g`)[1] + ((summary(firstTaxiAcc$`linear_g`)[6] - summary(firstTaxiAcc$`linear_g`)[1])*0.1)~ "1",
                   linear_g < summary(firstTaxiAcc$`linear_g`)[1] + ((summary(firstTaxiAcc$`linear_g`)[6] - summary(firstTaxiAcc$`linear_g`)[1])*0.2)~ "2",
                   linear_g < summary(firstTaxiAcc$`linear_g`)[1] + ((summary(firstTaxiAcc$`linear_g`)[6] - summary(firstTaxiAcc$`linear_g`)[1])*0.3)~ "3",
                   linear_g < summary(firstTaxiAcc$`linear_g`)[1] + ((summary(firstTaxiAcc$`linear_g`)[6] - summary(firstTaxiAcc$`linear_g`)[1])*0.4)~ "4",
                   linear_g < summary(firstTaxiAcc$`linear_g`)[1] + ((summary(firstTaxiAcc$`linear_g`)[6] - summary(firstTaxiAcc$`linear_g`)[1])*0.5)~ "5",
                   linear_g < summary(firstTaxiAcc$`linear_g`)[1] + ((summary(firstTaxiAcc$`linear_g`)[6] - summary(firstTaxiAcc$`linear_g`)[1])*0.6)~ "6",
                   linear_g < summary(firstTaxiAcc$`linear_g`)[1] + ((summary(firstTaxiAcc$`linear_g`)[6] - summary(firstTaxiAcc$`linear_g`)[1])*0.7)~ "7",
                   linear_g < summary(firstTaxiAcc$`linear_g`)[1] + ((summary(firstTaxiAcc$`linear_g`)[6] - summary(firstTaxiAcc$`linear_g`)[1])*0.8)~ "8",
                   linear_g < summary(firstTaxiAcc$`linear_g`)[1] + ((summary(firstTaxiAcc$`linear_g`)[6] - summary(firstTaxiAcc$`linear_g`)[1])*0.9)~ "9",
                   linear_g < summary(firstTaxiAcc$`linear_g`)[1] + ((summary(firstTaxiAcc$`linear_g`)[6] - summary(firstTaxiAcc$`linear_g`)[1])*1.0)~ "10")) %>% 
  arrange(desc(as.numeric(accelerationRisk)))
firstTaxiAccRisk
```
#getting the average risj rating for each taxi
```{r}
firstTaxiAccRiskAvg <- firstTaxiAccRisk %>% 
  group_by(vehicleid) %>% 
  drop_na(accelerationRisk) %>% 
  summarise(avgAccelerationRisk = mean(as.numeric(accelerationRisk)))
firstTaxiAccRiskAvg
```
#making one table of all the ratings for each taxi
```{r}
firstTaxiRisk <- firstTaxiSpeedRiskAvg %>% 
  left_join(firstTaxiCornerRiskAvg, firstTaxiSpeedRiskAvg, by = "vehicleid") %>% 
  left_join(firstTaxiBrakingRiskAvg, firstTaxiSpeedRiskAvg, by = "vehicleid") %>% 
  left_join(firstTaxiAccRiskAvg, firstTaxiSpeedRiskAvg, by = "vehicleid")
firstTaxiRisk
```
```{r}
saveRDS(firstTaxiRisk, file = "TaxiRiskAnalysis.rds") 
```

